Continuous time

### 2023-09-19 continuous-time-draft.qmd
styling
I think

```{r}
#| label: fig-duration
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
  filter(dat, 
    pid == "p6622",
    session %in% 14:16
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(prompt == "Wellbeing")),
    size = 2
    ) +
  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
  scale_y_continuous(breaks = 14:16) +
  facet_wrap(~pid)

p1 <- ggplot(
  filter(dat2, 
    pid == "p6622",
    session %in% 14:16
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(post)),
    size = 2
    ) +
  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
  scale_y_continuous(breaks = 14:16) +
  facet_wrap(~pid)
p0 / p1
```

@@

### 1. Session duration doesn't always start at 0
As can be seen in @fig-duration, session 16 isn't included in the pre/post data, even tho the first response is a WB prompt. This is because it's duration is 6. 
```{r}
filter(dat, 
  pid == "p6622",
  session %in% 14:16
) %>% 
kable()

filter(dat2, 
  pid == "p6622",
  session %in% 14:16
) %>% 
kable()
```

@@

```{r}
#| label: fig-pre-post
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
  filter(dat, 
    pid == "p10",
    session %in% 16:26
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(prompt == "Wellbeing")),
    size = 2
    ) +
  scale_x_datetime(
    limits = c(
      ymd_hms("2023-02-25 00:00:00"), 
      ymd_hms("2023-02-29 00:00:00"
      )
    )
  ) +
  scale_y_continuous(breaks = 16:26)

p1 <- ggplot(
  filter(dat2, 
    pid == "p10",
    session %in% 16:26
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(post)),
    size = 2
    ) +
  scale_x_datetime(
    limits = c(
      ymd_hms("2023-02-25 00:00:00"), 
      ymd_hms("2023-02-29 00:00:00")
      )
    ) +
  scale_y_continuous(breaks = 16:26)
p0 / p1
```

@@

```{r}
filter(dat, pid == "p10", session == 16) %>% 
  kable()

filter(dat2, pid == "p10", session == 16) %>% 
  kable()
```

@@

### Fit model
Reset duration within each session
```{r}
dat <- dat %>% 
  mutate(
    duration = duration - duration[1],
    hours = duration / 60,
    .by = c(pid, session)
  ) %>% 
  filter(prompt == "Wellbeing")
```

Fit linear change
```{r}
library(rms)
fit0 <- lmer(mood ~ hours + (1 | session) + (1 | pid), data = dat)
summary(fit0)
```

What's the distribution of session lengths
```{r}
median_ses_len <- dat %>% 
  group_by(pid, session) %>% 
  summarise(session_length = max(duration)) %>% 
  group_by(pid) %>% 
  summarise(median_len = median(session_length))

median(median_ses_len$median_len)

mean(median_ses_len$median_len > 60)

@@

"Typical" session length is 24 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.

```{r}
emm0 <- emmeans(
  fit0, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )

p0 <- as.data.frame(emm0) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Fit model with only random intercepts.
```{r}
fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 | pid), data = dat)
emm1 <- emmeans(
  fit1, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )
p1 <- as.data.frame(emm1) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Adding `rcs(hours, 5)` as random effects would be messy, so let's fit the model with a linear random slope at the participant level. I don't think random slopes at the session level is needed? Moreover, I would expect that the exact random slopes specification have little impact on fixed effects in this case.
```{r}
path <- "models/lmm-fit2.Rds"
if (file.exists(path)) {
  fit2 <- read_rds(path)
} else {
  fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 + hours | pid), data = dat)
  write_rds(fit2, path)
}
emm2 <- emmeans(
  fit2, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )
p2 <- as.data.frame(emm2) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

```{r}
p0 / p1
```

>>>>

```{r}
#| label: fig-duration
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
    filter(
        dat,
        pid == "p6622",
        session %in% 14:16
    ),
    aes(
        time,
        session,
        group = session
    )
) +
    geom_point(
        aes(color = factor(prompt == "Wellbeing")),
        size = 2
    ) +
    scale_x_datetime(
        limits = c(
            ymd_hms("2022-09-19 16:00:00"),
            ymd_hms("2022-09-20 00:00:00")
        )
    ) +
    scale_y_continuous(breaks = 14:16) +
    facet_wrap(~pid)

p1 <- ggplot(
    filter(
        dat2,
        pid == "p6622",
        session %in% 14:16
    ),
    aes(
        time,
        session,
        group = session
    )
) +
    geom_point(
        aes(color = factor(post)),
        size = 2
    ) +
    scale_x_datetime(
        limits = c(
            ymd_hms("2022-09-19 16:00:00"),
            ymd_hms("2022-09-20 00:00:00")
        )
    ) +
    scale_y_continuous(breaks = 14:16) +
    facet_wrap(~pid)
p0 / p1
```

@@

### 1. Session duration doesn't always start at 0
As can be seen in @fig-duration, session 16 isn't included in the pre/post data, even tho the first response is a WB prompt. This is because it's duration is 6. 
```{r}
filter(
    dat,
    pid == "p6622",
    session %in% 14:16
) |>
    kable()

filter(
    dat2,
    pid == "p6622",
    session %in% 14:16
) |>
    kable()
```

@@

```{r}
#| label: fig-pre-post
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
    filter(
        dat,
        pid == "p10",
        session %in% 16:26
    ),
    aes(
        time,
        session,
        group = session
    )
) +
    geom_point(
        aes(color = factor(prompt == "Wellbeing")),
        size = 2
    ) +
    scale_x_datetime(
        limits = c(
            ymd_hms("2023-02-25 00:00:00"),
            ymd_hms("2023-02-29 00:00:00")
        )
    ) +
    scale_y_continuous(breaks = 16:26)

p1 <- ggplot(
    filter(
        dat2,
        pid == "p10",
        session %in% 16:26
    ),
    aes(
        time,
        session,
        group = session
    )
) +
    geom_point(
        aes(color = factor(post)),
        size = 2
    ) +
    scale_x_datetime(
        limits = c(
            ymd_hms("2023-02-25 00:00:00"),
            ymd_hms("2023-02-29 00:00:00")
        )
    ) +
    scale_y_continuous(breaks = 16:26)
p0 / p1
```

@@

```{r}
filter(dat, pid == "p10", session == 16) |>
    kable()

filter(dat2, pid == "p10", session == 16) |>
    kable()
```

@@

### Session lengths

What's the distribution of session lengths
```{r}
median_ses_len <- dat |>
    group_by(pid, session) |>
    mutate(
        duration = duration - duration[1],
    ) |>
    summarise(session_length = max(duration)) |>
    group_by(pid) |>
    summarise(median_len = median(session_length))
median(median_ses_len$median_len)
mean(median_ses_len$median_len > 60)
```

"Typical" session length is 35 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.

### Fit model
Reset duration within each session
```{r}
dat <- dat |>
    mutate(
        ps = interaction(pid, session, sep = "_")
    ) |>
    filter(prompt == "Wellbeing") |>
    mutate(
        duration = duration - duration[1],
        hours = duration / 60,
        .by = c(pid, session)
    )
write_rds(dat, "data/ct-dev-data.rds")
```

Fit linear change
```{r}
#| code-fold: show
library(rms)
fit0 <- lmer(mood ~ hours + (1 | ps) + (1 | pid), data = dat)
```

```{r}
emm0 <- emmeans(
    fit0,
    ~hours,
    at = list(hours = seq(0, 2, length.out = 100)),
    lmer.df = "asymptotic"
)

p0 <- as.data.frame(emm0) |>
    ggplot(aes(hours, emmean)) +
    geom_line() +
    geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
    labs(x = "Session duration (hours)")
```

Fit model with only random intercepts.
```{r}
#| code-fold: show
fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 | pid), data = dat)
```

```{r}
emm1 <- emmeans(
    fit1,
    ~hours,
    at = list(hours = seq(0, 2, length.out = 100)),
    lmer.df = "asymptotic"
)

## for plotting knot placement
get_knot_coords <- function(fit) {
    knots <- attr(rcs(dat$hours, 5), "parms")
    knots_x <- rcs(knots, knots)
    b0 <- fixef(fit)[1]
    b <- fixef(fit)[-1]
    knots_y <- b0 + knots_x[, 1] * b[1] + knots_x[, 2] * b[2] + knots_x[, 3] * b[3] + knots_x[, 4] * b[4]
    data.frame(
        x = knots,
        y = c(knots_y)
    )
}
p1 <- as.data.frame(emm1) |>
    ggplot(aes(hours, emmean)) +
    geom_line() +
    geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
    geom_point(
        data = get_knot_coords(fit1),
        aes(x, y)
    ) +
    labs(x = "Session duration (hours)")
```

Adding `rcs(hours, 5)` as random effects would be messy, so let's fit the model with a linear random slope at the participant level. I don't think random slopes at the session level is needed? Moreover, I would expect that the exact random slopes specification have little impact on fixed effects in this case.

```{r}
#| code-fold: show
path <- "models/lmm-fit2.Rds"
if (file.exists(path)) {
    fit2 <- read_rds(path)
} else {
    fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | ps) + (1 + hours | pid), data = dat)
    write_rds(fit2, path)
}
```

```{r}
emm2 <- emmeans(
    fit2,
    ~hours,
    at = list(hours = seq(0, 2, length.out = 100)),
    lmer.df = "asymptotic"
)
p2 <- as.data.frame(emm2) |>
    ggplot(aes(hours, emmean)) +
    geom_line() +
    geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
    # geom_point(
    #   data = get_knot_coords(fit2),
    #   aes(x,y)
    # ) +
    labs(x = "Session duration (hours)")
```

```{r}
p0 / p1 / p2
```

## YOLO model with random slopes
Here I add a random slope to each piece of the spline, i.e. 4 random slopes at the participant level. The model is fit on a random subset of 1000 pids.
```{r}
#| cold-fold: show

# include 1000 participants
set.seed(1337)
pids <- sample(unique(dat$pid), 1000)
tmp <- filter(dat, pid %in% pids)
# so predict won't complain
tmp <- tmp |>
    mutate(
        pid = as.integer(pid),
        ps = as.integer(ps)
    )
# same spline but different parameterization
# usually less estimation problems
spline <- ns(
    dat$hours,
    knots = c(0.12, 0.4, 1),
    Boundary.knots = c(0.016, 3)
)
colnames(spline) <- paste0("hours", 0:3)
tmp <- cbind(dat, spline)
path <- "models/lmm-fit3.Rds"
if (file.exists(path)) {
    fit3 <- read_rds(path)
} else {
    fit3 <- lmer(
        mood ~ hours0 + hours1 + hours2 + hours3 +
         (1 | ps) + 
         (1 + hours0 + hours1 + hours2 + hours3 | pid),
        data = tmp
    )
    write_rds(fit3, path)
}

@@

I'm interested in if the spline changes based on the pre-session scores. Let's plot the fixed effects at -2, -1, 0 times the random intercept SD + the corresponding random slope values based on their correlation with the random intercept.

```{r}
get_trend <- function(fit, data, re_int = 0, re_s0 = 0, re_s1 = 0, re_s2 = 0, re_s3 = 0) {
    hours <- seq(0, 2, length.out = 100)
    x <- ns(
        hours,
        knots = c(0.12, 0.4, 1),
        Boundary.knots = c(0.016, 3)
    )
    b0 <- fixef(fit)[1] + re_int
    b <- fixef(fit)[-1]
    y <- b0 + x[, 1] * (b[1] + re_s0) + x[, 2] * (b[2] + re_s1) + x[, 3] * (b[3] + re_s2) + x[, 4] * (b[4] + re_s3)
    data.frame(
        x = hours,
        y = c(y),
        re_int
    )
}
re <- as.data.frame(VarCorr(fit3))
pid_int <- re[2, "sdcor"]
pid_slope0 <- re[3, "sdcor"]
pid_slope1 <- re[4, "sdcor"]
pid_slope2 <- re[5, "sdcor"]
pid_slope3 <- re[6, "sdcor"]
pid_cor0 <- re[7, "sdcor"]
pid_cor1 <- re[8, "sdcor"]
pid_cor2 <- re[9, "sdcor"]
pid_cor3 <- re[10, "sdcor"]
sd1 <- 1
sd2 <- 2
rbind(
    get_trend(fit3, data = tmp),
    get_trend(
        fit3,
        data = tmp,
        -sd1 * pid_int,
        -sd1 * pid_cor0 * pid_slope0,
        -sd1 * pid_cor1 * pid_slope1,
        -sd1 * pid_cor2 * pid_slope2,
        -sd1 * pid_cor3 * pid_slope3
    ),
    get_trend(
        fit3,
        data = tmp,
        -sd2 * pid_int,
        -sd2 * pid_cor0 * pid_slope0,
        -sd2 * pid_cor1 * pid_slope1,
        -sd2 * pid_cor2 * pid_slope2,
        -sd2 * pid_cor3 * pid_slope3
    )
) |>
    ggplot(
        aes(
            x,
            y,
            color = factor(re_int),
            group = factor(re_int)
        )
    ) +
    geom_line() +
    labs(x = "Session duration (hours)")
```

The random intercepts are only weakly correlated with each piece of the cubic spline, so the shape doesn't change much.

Let's also predict some individual slopes.

```{r}
pred <- filter(tmp, pid %in% sample(unique(tmp$pid), 10))
pred$yhat <- predict(fit3, newdata = pred)
ggplot(
    data = pred,
    aes(
        x = hours,
        y = yhat,
        group = ps
    )
) +
    geom_line(
        color = "blue",
        linewidth = 1
    ) +
    geom_point(
        aes(y = mood)
    ) +
    facet_wrap(~pid)
```

Zoom in on pid 9263.
```{r}
ggplot(
    data = filter(pred, pid == 9263),
    aes(
        x = hours,
        y = yhat,
        group = ps
    )
) +
    geom_line(
        aes(y = mood),
        linetype = "dashed",
        alpha = 0.33
    ) +
    geom_line(
        color = "blue",
        linewidth = 2
    ) +
    geom_point(
        aes(y = mood)
    ) +
    facet_wrap(~session)
```

Probably good enough, even tho the session effect is only modelled using a random intercept.

### Predict change
Another YOLO approach to estimate individual change during a session is to to look at the model predictions.

TODO: should be based on the actual session length, not only the range of WB prompts.

```{r}
pred <- fit3@frame
pred$yhat <- predict(fit3)
pred |>
    # compare the last and first value
    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
    # average over sessions within participants
    dplyr::summarize(change = mean(change), .by = pid) |>
    # average over participants
    dplyr::summarize(mean(change), sd(change))
```

We can also plot the distribution of the participants average change during sessions.
```{r}
pred |>
    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
    dplyr::summarize(change = mean(change), .by = pid) |>
    ggplot(aes(change)) +
    geom_histogram()
```

They really don't change much. Let's look at a participant out in the tails.
```{r}
pred |>
    dplyr::summarize(change = last(yhat) - yhat[1], .by = c(pid, ps)) |>
    filter(change > 0.1) |>
    head()
```

Zoom in on pid 23.
```{r}
pred2 <- filter(tmp, pid == 23)
pred2$yhat <- predict(fit3, newdata = pred2)
ggplot(
    data = pred2,
    aes(
        x = hours,
        y = yhat,
        group = ps
    )
) +
    geom_line(
        aes(y = mood),
        linetype = "dashed",
        alpha = 0.33
    ) +
    geom_line(
        color = "blue",
        linewidth = 2
    ) +
    geom_point(
        aes(y = mood)
    ) +
    facet_wrap(~session)
```

This looks kinda reasonable?

## brms
Ignore for now.

```{r}
#| code-fold: show
library(brms)
options(
    mc.cores = 8,
    brms.backend = "cmdstanr"
)
fit_b0 <- brm(
    mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 | pid),
    data = tmp,
    #threads = 2,
    chains = 4,
    iter = 500,
    backend = "rstan",
    file = "models/brm-pre-post-ct0"
)
pp_check(fit_b0)
```

Censoring
```{r}
#| code-fold: show
tmp <- tmp |>
    mutate(
        cl = case_when(
            mood == 0 ~ "left",
            mood == 1 ~ "right",
            TRUE ~ "none"
        )
    )
fit_b1 <- brm(
    mood | cens(cl) ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 | pid),
    data = tmp,
    threads = 2,
    iter = 1000,
    file = "models/brm-pre-post-ct1"
)
pp_check(fit_b1)
```

## GAM(M)

Let's try some GAMs. 
```{r}
library(mgcv)
knots <- list(hours = c(0.016, 0.12, 0.4, 1, 3))
tmp <- tmp |>
    mutate(
        pid = as.factor(pid),
        ps = as.factor(ps)
    )
```

We'll start with a cubic spline using the same knot placement as used with the LMMs.
```{r}
#| code-fold: show
fit_g_0 <- gam(
    mood ~ s(hours, bs = "cr", k = 5),
    knots = knots,
    method = "REML",
    data = tmp,
)
fit_g_0 <- gam(
    mood ~ s(hours, k = 15),
    #knots = knots,
    method = "REML",
    data = filter(tmp, hours < 5),
)
```

```{r}
plot_gam <- function(fit, t_end = 2) {
    emm0 <- emmeans(
        fit,
        ~hours,
        at = list(hours = seq(0, t_end, length.out = 100))
    )
    p0 <- as.data.frame(emm0) |>
        ggplot(aes(hours, emmean)) +
        geom_line() +
        geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
        labs(x = "Session duration (hours)")
    p0
}
plot_gam(fit_g_0) + lims(y = c(0,1))
```

we could add a random intercept for pid and session like this
```{r}
#| eval: false
#| code-fold: show
fit_g_1 <- gam(
    mood ~ s(hours, bs = "cr", k = 5) + s(pid, bs = "re") + s(ps, bs = "re"),
    knots = knots,
    data = tmp,
)
```

The call above is super slow, so let's use `gamm4` instead.

```{r}
#| code-fold: show
library(gamm4)
fit_g_1_lme4 <- gamm4(
    mood ~ s(hours, bs = "cr", k = 5),
    knots = knots,
    data = tmp,
    random = ~ (1 | pid) + (1 | ps)
)
```

```{r}
pred <- data.frame(hours = seq(0, 2, length.out = 100))
pred$yhat <- predict(fit_g_1_lme4$gam, newdata = pred)
ggplot(
    pred,
    aes(hours, yhat)
) +
    geom_line() +
    lims(y = c(0,1)) +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)")
```

We are using the same smooth for everyone here, which is probably not what we want.

### Separate smooth per participant
Let's try using a separate smooth per pid. We only fit the model to a subset of participants to save time.

```{r}
#| code-fold: show
set.seed(6666)
tmp2 <- filter(tmp, pid %in% sample(unique(pid), 200)) |>
    filter(!is.na(mood), hours <= 10)
fit_g_by_lme4 <- gamm4(
    mood ~ s(hours, by = pid, bs = "cr", k = 5),
    knots = knots,
    data = tmp2,
    random = ~ (1 | pid) + (1 | ps)
)
```

N.B. that only the GAM part is plotted, I'm not sure if there is a command that automagically includes the random effects when using `gamm4`.

```{r}
pred <- expand.grid(
    hours = seq(0, 2, length.out = 100),
    pid = unique(tmp2$pid)
)
pred$yhat <- predict(fit_g_by_lme4$gam, newdata = pred)
pred$model <- "by"
ggplot(
    pred,
    aes(
        hours,
        yhat,
        group = pid,
        color = pid
    )
) +
    geom_line() +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)") +
    theme(legend.position = "none")

plot_pids <- sample(unique(tmp2$pid), 10)
tmp2 <- filter(tmp2, !is.na(mood))
tmp2$yhat <- predict(fit_g_by_lme4$gam)
tmp2$model <- "by"
ggplot(
    filter(tmp2, pid %in% plot_pids),
    aes(
        hours,
        yhat,
        group = ps,
        color = pid
    )
) +
    geom_line(aes(y = mood)) +
    geom_line(color = "blue", linewidth = 1) +
    geom_point(aes(y = mood)) +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)") +
    facet_wrap(~pid) +
    theme(legend.position = "none")

pred_all <- filter(tmp2, pid %in% plot_pids)
```

We're fitting a separate fixed and random effects for each pid. Fitting this to the full data will be super slow?
```{r}
summary(fit_g_by_lme4$mer)
```

### Factor-smooth interaction 
We can also try a factor-smooth interaction instead, this should be closer to what we want.

```{r}
#| code-fold: show

# TODO: include a main effect
fit_g_fs_lme4 <- gamm4(
    mood ~ s(hours, bs = "cr", k = 5) + s(hours, pid, bs = "fs", xt = list(bs = "cr"), k = 5),
    knots = knots,
    data = tmp2,
    random = ~ (1 | pid) + (1 | ps)
)
plot(fit_g_fs_lme4$gam, select = 1)
```

`gamm4` is not supported by `emmeans`, and I'm not sure how to plot the "marginal" smooth.

```{r}
pred <- expand.grid(
    hours = seq(0, 2, length.out = 100),
    pid = unique(tmp2$pid)
)
pred$yhat <- predict(fit_g_fs_lme4$gam, newdata = pred)
pred$model <- "fs"
ggplot(
    pred,
    aes(
        hours,
        yhat,
        group = pid,
        color = pid
    )
) +
    geom_line() +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)") +
    theme(legend.position = "none")

tmp2$yhat <- predict(fit_g_fs_lme4$gam)
tmp2$model <- "fs"
ggplot(
    filter(tmp2, pid %in% plot_pids),
    aes(
        hours,
        yhat,
        group = ps,
        color = pid
    )
) +
    geom_line(aes(y = mood)) +
    geom_line(color = "blue", linewidth = 1) +
    geom_point(aes(y = mood)) +
    lims(y = c(0, 1)) +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)") +
    facet_wrap(~pid) +
    theme(legend.position = "none")
pred_all <- rbind(pred_all, filter(tmp2, pid %in% plot_pids))
```

The smooths in the plot above are all linear. Although, it looks pretty reasonable compared to the observed data. (TODO: investigate with more data.) If we fit a simple smooth to the data it's also linear.

```{r}
#| code-fold: show
gam(
    mood ~ s(hours, bs = "cr", k = 5),
    knots = knots,
    method = "REML",
    data = tmp2,
) |>
    plot_gam()

### LMM again
As a comparisons, let's refit the LMM with a cubic spline to the subset of data used here. And then we predict without including the random intercepts.

```{r}
#| code-fold: show
fit3 <- lmer(
  mood ~ hours0 + hours1 + hours2 + hours3 + (1 | ps) + (1 + hours0 + hours1 + hours2 + hours3 | pid), 
  data = tmp2
)
```

```{r}
tmp2$yhat <- predict(
    fit3,
    re.form = ~ (0 + hours0 + hours1 + hours2 + hours3 | pid)
)
tmp2$model <- "lmer"
ggplot(
    filter(tmp2, pid %in% plot_pids),
    aes(
        hours,
        yhat,
        group = ps,
        color = pid
    )
) +
    geom_line(aes(y = mood)) +
    geom_line(color = "blue", linewidth = 1) +
    geom_point(aes(y = mood)) +
    lims(y = c(0, 1)) +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)") +
    facet_wrap(~pid)

pred_all <- rbind(pred_all, filter(tmp2, pid %in% plot_pids))
```

### Combined plot
Plot the predictions from the 3 models in a combined figure.

```{r}
#| fig.height: 5
#| fig.width: 7
ggplot(
    pred_all,
    aes(
        hours,
        yhat,
        group = interaction(ps, model),
        color = model
    )
) +
    geom_line(data = filter(pred_all, model == "fs"), aes(y = mood, color = NA), alpha = 0.5) +
    geom_point(data = filter(pred_all, model == "fs"), aes(y = mood, color = NA), alpha = 0.5) +
    geom_line(linewidth = 1.5) +
    # lims(y = c(0, 1)) +
    # geom_ribbon(aes(ymin = lower.CL, ymax = upper.CL), alpha = 0.25) +
    labs(x = "Session duration (hours)") +
    facet_wrap(~pid, scales = "free_y")
```

### GAM summary
Assuming that we wanted to fit these smooths in `brms` to model censoring, then

- separate smooths using `by` is not feasible?
- factor-smooth interaction is maybe feasible using the full data.
  - Might be overpenalizing the smooths as they were all straight lines? 
- the LMM is basically in between the model above when it come to "wiggliness", as would be expected

I'm not sure if some if the issues is caused by how GAMs are affected by the highly irregular sessions + missing data.

LMM might be _the_ model?
